Generating Images from Captions with Attention

### 2015-10-09 ram_workshop/ram-paper.tex
update

\vspace*{-1.0cm}
\title{Generating Images from Captions with Attention}

@@

Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image. Our proposed model 
defines a generative process of images conditioned on the caption. In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$. Our model can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}.

\vspace{-0.1in}
\subsection{Language Representation: the Bidirectional Attention RNN}
\label{sec:lang}
We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN. In a Bidirectional RNN, the two LSTMs process the input sequence from both forward and backward directions. They produce the hidden states sequences $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ and $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$ respectively. These hidden states are then concatenated together into the final sentence representation $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.

@@

\begin{figure}[!h]
\vspace{-0.3in}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/alignDrawNew.pdf}\quad
%
\end{center}
\vspace{-0.2in}
\caption{alignDRAW: Generative model of images conditioned on captions.}
\label{fig:figmodel}
\vspace{-0.2in}
\end{figure}
%\vspace{-0.1in}

@@

% \vspace{-0.5cm}
\begin{figure}[!h]
 \vspace{-0.5cm}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A \underline{yellow} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{red} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{green} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{blue} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the color while keeping the caption fixed.}
\label{fig:genimages1}
\vspace{-0.6cm}
\end{figure}

Apart from changing the colors of objects, we experimented with changing the background of the scene described in a caption to see whether this would result in the appropriate changes in the generated samples. The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area. Nevertheless, as shown in Figure~\ref{fig:genimages2}, changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image. The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.

\vspace{-0.5cm}
\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\\
\vspace{-0.45cm}
%
\subfloat[A very large commercial plane flying in \underline{blue} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A very large commercial plane flying in \underline{rainy} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{dry} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{green} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
%
\end{center}
\caption{{\bf Bottom}: Examples of changing the background while keeping the caption fixed. {\bf Top}: The respective nearest training images based on pixel-wise L2 distance.}
\label{fig:genimages2}
\vspace{-0.4cm}
\end{figure}

@@

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
%
\subfloat[\underline{A bowl of bananas} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{cat}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{dog}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the object while keeping the caption fixed.}
\label{fig:genimages3}
\vspace{-0.3cm}
\end{figure}

@@

\vspace{-0.5cm}
\begin{figure}[!h]
\vspace{-0.5cm}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{desert}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
%
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{forest}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
%
\subfloat[\hlctwo{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlctwo{beach}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
%
\subfloat[\hlcthree{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlczero{sun}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
%
\end{center}
\caption{Examples of most attended words while changing the background in the caption.}
\label{fig:attentimages1}
\vspace{-0.4cm}
\end{figure}

@@

\vspace{-0.5cm}
\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
%
\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
%
\subfloat[Conv. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
%
\subfloat[Fully-conn. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
%
\end{center}
\caption{Four different models displaying results from sampling caption ``A group of people walk on a beach with surf boards". LAPGAN samples look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect.}
\label{fig:diffmodels}
\vspace{-0.4cm}
\end{figure}

@@

In addition we calculate Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation. The metric is also calculated on small windows of the image. We sampled 50 images from the prior of each generative model for every test caption and calculated SSI, which is reported in Table~\ref{tab:results}.

\begin{table}[!h]
\small
\begin{center}
\begin{tabulary}{\linewidth}{c || c c c c c || c}
%\hline
%\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
%\hline
& \multicolumn{5}{c||}{Image Search} & Image Similarity \\
%\cline{2-6}
\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
\hline
\hline
LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
\hline
Fully-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
noalignDRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
alignDRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
\end{tabulary}
\end{center}
\vspace{-0.1cm}
\caption{Results of different models on COCO dataset (before sharpening).}
\label{tab:results}
\vspace{-0.3cm}
\end{table}

{\footnotesize
\bibliography{ram-paper}}

>>>>

%\vspace*{-1.0cm}
\title{Generating Images from Captions with Attention}

@@

Our proposed model 
defines a generative process of images conditioned on the caption. In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$. Our model can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}.

\vspace{-0.1in}
\subsection{Language Representation: the Bidirectional Attention RNN}
\label{sec:lang}
Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image. We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN. In a Bidirectional RNN, the two LSTMs process the input sequence from both forward and backward directions. They produce the hidden states sequences $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ and $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$ respectively. These hidden states are then concatenated together into the final sentence representation $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.

@@

\begin{figure}[!h]
\vspace{-0.3in}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/alignDrawAnnotated.pdf}\quad
%
\end{center}
\vspace{-0.2in}
\caption{alignDRAW: Generative model of images conditioned on captions.}
\label{fig:figmodel}
\vspace{-0.2in}
\end{figure}
%\vspace{-0.1in}

@@

% \vspace{-0.5cm}
\begin{figure}[!h]
 \vspace{-0.5cm}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[A \underline{yellow} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{red} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{green} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{blue} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{Examples of changing the color while keeping the caption fixed.}
\label{fig:genimages1}
\vspace{-0.3cm}
\end{figure}
Apart from changing the colors of objects, we experimented with changing the background of the scene described in a caption to see whether this would result in the appropriate changes in the generated samples. The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area. Nevertheless, as shown in Figure~\ref{fig:genimages2}, changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image. The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.

\vspace{-0.5cm}
\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\\
\vspace{-0.45cm}
%
\subfloat[A very large commercial plane flying in \underline{blue} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A very large commercial plane flying in \underline{rainy} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{dry} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{green} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{{\bf Bottom}: Examples of changing the background while keeping the caption fixed. {\bf Top}: The respective nearest training images based on pixel-wise L2 distance.}
\label{fig:genimages2}
\vspace{-0.3cm}
\end{figure}

@@

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
%
\subfloat[\underline{A bowl of bananas} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{cat}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{dog}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{Examples of changing the object while keeping the caption fixed.}
\label{fig:genimages3}
\vspace{-0.3cm}
\end{figure}

@@

%\vspace{-0.3cm}
\begin{figure}[!h]
\vspace{-0.5cm}
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{desert}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
%
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{forest}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
%
\subfloat[\hlctwo{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlctwo{beach}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
%
\subfloat[\hlcthree{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlczero{sun}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{Examples of most attended words while changing the background in the caption.}
\label{fig:attentimages1}
\vspace{-0.4cm}
\end{figure}

@@

\vspace{-0.5cm}
\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
%
\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
%
\subfloat[Conv. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
%
\subfloat[Fully-conn. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
%
\end{center}
\vspace{-0.3cm}
\caption{alignDRAW and three baseline models conditioned on skip-thoughts \cite{skip-thought} displaying results from sampling caption ``A group of people walk on a beach with surf boards". LAPGAN samples look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect.}
\label{fig:diffmodels}
\vspace{-0.3cm}
\end{figure}

@@

In addition we calculate Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation. The metric is also calculated on small windows of the image. We sampled 50 images from the prior of each generative model for every test caption and calculated SSI, which is reported in Table~\ref{tab:results}.
\vspace{-0.12in}
\begin{table}[!h]
\small
\begin{center}
\begin{tabulary}{\linewidth}{c || c c c c c || c}
%\hline
%\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
%\hline
& \multicolumn{5}{c||}{Image Search} & Image Similarity \\
%\cline{2-6}
\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
\hline
\hline
LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
\hline
Fully-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
noalignDRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
alignDRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
\end{tabulary}
\end{center}
\vspace{-0.1cm}
\caption{Results of different models on COCO dataset (before sharpening).}
\label{tab:results}
\vspace{-0.3cm}
\end{table}
\vspace{-0.3in}
{\footnotesize
\bibliography{ram-paper}}
